---
title: "Albundy Shoe Sales"
output:
  pdf_document: default
  html_document: default
---
Objective: Compute confidence intervals for men shoe sales in the US between 2015 and 2016.

Goal: Determine how many shoes of each size should be in stock each month of the year to ensure customers will have product to buy and avoid overstocking.
```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

```{r}
getwd() # check working directory
setwd("/Users/mine/Downloads") #change to location of file to be imported
getwd() #check that file index has been changed
```


```{r}
library("readxl") #load required library to read excel file
shoe_data <- read_excel("3.17.Practical-example.Confidence-intervals-exercise.xlsx")
#read in the file


```


```{r}
shoe_data <- shoe_data[-c(1,2),]
#Remove empty rows in beginning of file
names(shoe_data) <- shoe_data[1,] 
#Assign column names to the lead row
shoe_data <- shoe_data[-1,] 
# Remove lead row as it has now replaced the column names
colnames(shoe_data)[12]<- "Year" 
# assign column name to NA column after determining best fit column name
```


```{r}
library("openxlsx") 
# load required library to adjust date column
shoe_data$Date <- convertToDate(shoe_data$Date) 
# function to convert excel dates into readable dates in R

```


```{r}
us_data_2016 <- subset(shoe_data, Country !="United States" 
                       & Year == "2016"& Gender == "Male") 
#Subset data for US only, year of 2016, and gender is male
us_data_2015 <- subset(shoe_data, Country !="United States" 
                       & Year == "2015"& Gender == "Male") 
#This is done as our goal is to compare the sales between the two years.
```


```{r}

us_data_2015$month<-months(us_data_2015$Date) 
# Insert month column  to create a table of size sales by month
us_data_2016$month<-months(us_data_2016$Date)
t_2015<- table(as.numeric(us_data_2015$`Size (US)`),
               factor(us_data_2015$month,levels=month.name))
# create a table with the sizes and months for each year
t_2016<- table(as.numeric(us_data_2016$`Size (US)`),
               factor(us_data_2016$month,levels=month.name))
```


```{r}

df_2015<-as.data.frame.matrix(t_2015) 
#Convert tables to dataframes for easier data calculations
df_2016<-as.data.frame.matrix(t_2016) 
```


```{r}
df_combined<-cbind(df_2015, df_2016)
# combine two dataframes 2015 first 2016 second


```


```{r}
library("matrixStats")#load library to compute row operations on dataframes
mean<- rowMeans(df_combined) #Yields means by row
std_error<- rowSds(as.matrix(df_combined))/sqrt(24) 
# yields standard deviation by row and then computes standard error

margin_error<- std_error*qt(.025, 23, lower.tail=FALSE)


```


```{r}
lower_ci<-mean-margin_error 
# calculate upper and lower CI for mens shoe sales in the US
upper_ci<-mean+margin_error


```


```{r}
df_combined$mean<-mean # adding new calculations to existing dataframe
df_combined$std_error<-std_error
df_combined$margin_error<-margin_error
df_combined$lower_ci<-lower_ci
df_combined$upper_ci<-upper_ci
df_combined$pairstocarry<-round(upper_ci)


```

```{r}
df_combined[,c(25,30)]


```
In conclusion the pairstocarry column represents the rough number of shoes to have in stock per size each month of the year.
With this information I can be 95% percent sure that the number of shoes purchased will not exceed the value in the pairstocarry column for any given month of the year.

